---
title: Semester Schedule Planner
author: Derek Sollberger
date: '2018-08-01'
slug: semester-schedule-planner
categories:
  - teaching
tags:
  - teaching
  - xts
  - timeSeries
header:
  caption: ''
  image: ''
---

As a lecturer at UC Merced, each semester I find myself needing to figure out how many days I get to teach.  Furthermore, I like to plan out my lecture schedule down to the days themselves.  Here I hope to use some `R` code to quickly figure out (say) which days are on Tuesday or Thursday.

## Construction

In `R`, the `seq` command can be used with dates easily.

```{r}
semester_dates <- seq(as.Date("2018-08-22"), as.Date("2018-12-07"), by = "days")
```

When working with time data, the `xts` package ("extensible time series") offers convenient tools for manipulating time series.

```{r, message = FALSE, warning = FALSE}
library("xts")
xts_dates <- as.xts( rep(NA, length(semester_dates)), order.by = semester_dates)
```

The convention is that "0" is a Sunday, "1" is a Monday, ..., and "5" is a Friday.

## Monday and Wednesday

```{r}
Mondays <- xts_dates[.indexwday(xts_dates) == 1]
head(Mondays)

MW_dates <- xts_dates[.indexwday(xts_dates) %in% c(1, 3)]
head(MW_dates)
```

## Tuesday and Thursday

```{r}
TuTh_dates <- xts_dates[.indexwday(xts_dates) %in% c(2, 4)]
head(TuTh_dates)
```

## Friday

```{r}
Fri_dates <- xts_dates[.indexwday(xts_dates) == 5]
head(Fri_dates)
```

## Exporting the Time Series

From here, I might still want to use Excel for data entry in my planning.  `xts` objects are still `zoo` objects.

```{r}
#write.zoo(Mondays, sep = ",", file = "schedule.csv")
#write.zoo(MW_dates, sep = ",", file = "schedule.csv")
#write.zoo(TuTh_dates, sep = ",", file = "schedule.csv")
write.zoo(Fri_dates, sep = ",", file = "schedule.csv")
```

Now I have a spreadsheet with the dates already arranged!