---
title: Supreme Court Confirmations (1967-present)
author: Derek Sollberger
date: '2018-10-07'
slug: supreme-court-confirmations-1967-present
categories:
  - politics
tags:
  - ggplot
  - plotly
header:
  caption: ''
  image: ''
---

### Introduction

Following up on [Rachel Wellford's tweet](https://twitter.com/rachelwellford/status/1048308997219082240) about Senate votes for Supreme Court confirmations, I decided to try to graph the data.  Below, I have 

* a `ggplot` picture with decent labeling
* a searchable `datatable`
* a `plotly` interactive graph

The data came from [Senate.gov](https://www.senate.gov/pagelayout/reference/nominations/Nominations.htm).  I chose to focus on 1967 onward because it appeared that voting procedures were slightly different before Thurgood Marshall's nomination process.  Otherwise, I would have chosen to go back to when the Senate first had 100 members.

```{r, message = FALSE, warning = FALSE}
library("DT")
library("ggplot2")
library("ggrepel")
library("plotly")
library("readxl")
scc <- read_excel("../../static/Supreme Court confirmations.xlsx")
```

### ggplot

```{r}
ggplot(scc, aes(x = Nay, y = Yea, label = Last)) +
  geom_point() +
  geom_text_repel() +
  labs(x = "Nay votes", y = "Yea votes", 
       title = "Supreme Court Confirmations (1967 to present)",
       subtitle = "Marshall to Kavanaugh",
       caption = "Source: Senate.gov") + 
  xlim(c(-10, 60)) +
  ylim(c(40, 110))
```

### Datatable

```{r}
DT::datatable(scc)
```

### Plotly

Here is a slightly interactive graph built with a combination of `ggplot` and `plotly`:

```{r}
basemap <- ggplot(scc, aes(x = Nay, y = Yea, label = Last)) +
  geom_point() +
  geom_text(nudge_x = 5) + # geom_text_repel is not supported by Plotly yet
  labs(x = "Nay votes", y = "Yea votes", 
       title = "Supreme Court Confirmations (1967 to present)",
       subtitle = "Marshall to Kavanaugh",
       caption = "Source: Senate.gov") + 
  xlim(c(-10, 60)) +
  ylim(c(40, 110))
ggplotly(basemap)
```

Here is a slightly interactive graph with readable labels:

```{r}
t <- list(
  family = "sans serif",
  size = 14,
  color = toRGB("grey80"))

plot_ly(scc, x = ~Nay, y = ~Yea, text = ~Nominee) %>%
  add_markers() %>%
  add_text(textfont = t, textposition = "top right")
```

Finally, here is a fully interactive graph, but the labels are not revealed until someone hovers their mouse cursor over the points:

```{r}
plot_ly(scc, x = ~Nay, y = ~Yea, type = 'scatter', mode = 'markers',
        hoverinfo = 'text',
        text = ~paste('<b>Nominee:</b> ', Nominee,
                      '<br> Yea: ', Yea,
                      '<br> Nay: ', Nay,
                      '<br> Nominated: ', Nominated,
                      '<br> Confirmed: ', Confirmed,
                      '<br> President: ', President))
```

