---
title: Curse of Dimensionality
author: Derek Sollberger
date: '2019-01-08'
slug: curse-of-dimensionality
categories:
  - linear algebra
tags:
  - math
header:
  caption: ''
  image: ''
---

Today, I hope to present a quick glimpse at the phenomenon called the "Curse of Dimensionality".  For this demonstration, I am simply calculating how much random data stays within two standard deviations (in the Euclidean norm) as we go from one dimension to higher dimensions.

### Random Data

Here are 10 vectors of 100 random numbers each sampled from the standard normal distribution stored as a matrix ...

```{r}
X <- matrix(rnorm(1000), nrow = 100, ncol = 10)
```

... and as a data frame.

```{r}
df <- data.frame(X)
head(df)
```

### One Dimension

For normally distributed data, we expect that about 95% of data falls within two standard deviations.

```{r}
x1 <- X[,1]
within2sd <- abs(x1) <= 2
df1 <- data.frame(x1, within2sd)
mean(within2sd)
```

In this example, `r round(mean(within2sd)*100, 2)` percent of the data in the first vector is within two standard deviations of the mean.

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
df1 %>% 
  ggplot(aes(x = x1, , y = 0, color = within2sd)) +
  geom_point() + 
  labs(title = "One Dimension of Normal Distribution Data")
```

### Two Dimensions

However, when we go into two or more dimensions, the colloquial "95%" expection starts to fade.  To aid calculations, the `row_norms` function in the `slam` package uses the Euclidean norm by default.  To aid visualization, we will use a helper function (found on Stack Overflow at https://stackoverflow.com/questions/6862742/draw-a-circle-with-ggplot2) to draw one circle.

```{r, message = FALSE, warning = FALSE}
library(slam)
within2sd <- row_norms(X[,1:2]) <= 2

gg_circle <- function(r = 1, xc = 0, yc = 0, color="black", fill=NA, ...) {
    x <- xc + r*cos(seq(0, pi, length.out=100))
    ymax <- yc + r*sin(seq(0, pi, length.out=100))
    ymin <- yc + r*sin(seq(0, -pi, length.out=100))
    annotate("ribbon", x=x, ymin=ymin, ymax=ymax, color=color, fill=fill, ...)
}

df2 <- data.frame(X[,1:2], within2sd)
df2 %>%
  ggplot(aes(x = X1, y = X2, color = within2sd)) +
  geom_point() + 
  gg_circle(r = 2, color = "red") +
  coord_fixed() + 
  labs(title = "Two Dimensions of Normal Distribution Data")

mean(within2sd)
```

In this example, `r round(mean(within2sd)*100, 2)` percent of the data in the first 2 vectors is within two standard deviations of the mean.

### Higher Dimensions

Plotting scatterplots in higher dimensions is much more complicated, but we can still perform the `norm` calculations pretty quickly.

```{r}
N <- 10 #total number of dimensions
within2sd <- rep(0, N) #initialization

# one dimension
within2sd[1] <- mean(abs(x1) <= 2)

# higher dimensions
for(d in 2:N){
  within2sd[d] <- mean(row_norms(X[,1:d]) <= 2)
}

# plot
dimensions <- 1:N
df <- data.frame(dimensions, within2sd)
df %>%
  ggplot(aes(x = dimensions, y = within2sd)) +
  geom_bar(stat = "identity", fill = "blue") +
  scale_x_continuous("Dimensions", breaks = 1:N)

```

