---
title: Graduation Rates
author: Derek Sollberger
date: '2019-01-08'
slug: graduation-rates
categories:
  - stochastic processes
tags:
  - math
header:
  caption: ''
  image: ''
---

In this post, I want to run an example of absorbing states in stochastic processes.  This example is based on Example 3.29 in *Introduction to Stochastic Processes* in R by Robert Dobrow.

# Data

The data I have assembled is based on [IRDS reports](https://irds.ucmerced.edu/student-data) from my own University of California at Merced.

```{r}
# weights
weights <- c(20, 70, 0, 0, 10, 0, 
       0, 5, 89, 0, 6, 0, 
       0, 0, 3, 94, 3, 0, 
       0, 0, 0, 24, 1, 76, 
       0, 0, 0, 0, 100, 0, 
       0, 0, 0, 0, 0, 100)
sparse_weights <- weights[weights > 0]

# transition matrix (row stochastic)
P <- matrix(weights, nrow = 6, byrow = TRUE)*0.01

# states
rownames(P) <- c("Freshman", "Sophomore", "Junior", "Senior", "Dropped", "Graduated")
colnames(P) <- c("Freshman", "Sophomore", "Junior", "Senior", "Dropped", "Graduated")
P
```

This formulation assumes that about 10 percent of students drop after the first year, and about 24 percent of seniors continue to take classes into the fifth year or more.

# Network Graph

```{r, message = FALSE, warning = FALSE}
library(igraph)
network <- graph_from_adjacency_matrix(P, mode = "directed", weighted = TRUE)
plot(network,
     edge.arrow.size = 0.5,
     edge.color = "#D4AE00",
     edge.label = sparse_weights,
     layout = layout_in_circle(network))
```


# Absorbing States

```{r}
# canonical form
Q <- P[1:4, 1:4]
R <- P[1:4, 5:6]

# fundamental matrix (inverse of (I - Q))
fundamental_matrix <- solve(diag(4) - Q)

# absorption
absorption <- fundamental_matrix %*% R
absorption
```

This model predicts that, starting with incoming students, about 79 percent of those students will eventually graduate while about 21 percent will eventually drop from UC Merced enrollment.