---
title: Valentine's Day 2019
author: Derek Sollberger
date: '2019-02-14'
slug: valentine-s-day-2019
categories:
  - gganimate
tags:
  - gganimate
  - heart
header:
  caption: ''
  image: ''
  preview: yes
---

In this short project, I hope to draw a heart through an animation of appearing dots.

```{r, message = FALSE, warning = FALSE}
library("gganimate")
library("ggforce")
library("tidyverse")
```

# Draw a Heart

Here I will place two circles centered at $(\pm 1, 1)$ with the same radius $r = \sqrt{2}$.

```{r}
circles <- data.frame(
  x0 = c(-1,1),
  y0 = rep(1,2),
  r = rep(1, 2)
)

right_pt <- (sqrt(2) + 1) / sqrt(2)
left_pt <-  -1*right_pt
f <- function(x){abs(x) - sqrt(2)}

ggplot(data.frame(x = c(left_pt, right_pt)), aes(x)) +
  coord_fixed() +
  geom_circle(aes(x0 = x0, y0 = y0, r = r), data = circles, inherit.aes = FALSE) +
  stat_function(fun = f, geom = "line")
```

# Generate Some Dots

Now let's see if I can get some red dots into a heart shape and green dots outside of the heart.

```{r}
# initialization
N <- 10000
x <- runif(N, -2, 2)
y <- runif(N, -sqrt(2), 2)
colors = c("white", "red", rep("green", N-2)) #factor levels set early
state = rep(1, N)

# separate colors
for (i in 4:N){
  if (y[i] < 1){
    if (y[i] > (abs(x[i]) - sqrt(2))){colors[i] = "red"}
  } else {
    if(sqrt((x[i] - 1)^2 + (y[i] - 1)^2) < 1 | 
       sqrt((x[i] + 1)^2 + (y[i] - 1)^2) < 1){colors[i] = "red"}
  }
}

dot_df <- data.frame(x,y, colors, state)
dot_df %>%
  ggplot(aes(x = x, y = y)) +
    coord_fixed() +
    geom_point(color = colors) +
    theme_minimal()
```

# Animation

My initial plan was to have the dots appear one at a time to gradually make the heart, but the data storage was prohibitive.  This back up plan was to literally make a new set of dots for each animation frame---and color the dots again!---and it comes with a neat "heartbeat" effect.  The render time for the animated GIF was about 2 minutes on my computer.

```{r}
max_iter <- 14 #number of animation frames

for(j in 1:max_iter){
  x <- runif(N, -2, 2)
  y <- runif(N, -sqrt(2), 2)
  colors = c("white", "red", rep("green", N-2)) #factor levels set early
  
  for (i in 4:N){
    if (y[i] < 1){
      if (y[i] > (abs(x[i]) - sqrt(2))){colors[i] = "red"}
    } else {
      if(sqrt((x[i] - 1)^2 + (y[i] - 1)^2) < 1 | 
         sqrt((x[i] + 1)^2 + (y[i] - 1)^2) < 1){colors[i] = "red"}
    }
  }

  state <- rep(j, N)
  this_df <- data.frame(x, y, colors, state)
  #colnames(this_df) <- c("x", "y", "colors", "state")
  dot_df <- rbind(dot_df, this_df)
}

dot_df %>%
  ggplot(aes(x = x, y = y)) +
    coord_fixed() +
    geom_point(color = dot_df$colors) +
    labs(title = "Valentine's Day 2019",
         subtitle = "Frame: {closest_state}",
         caption = "Made by Derek Sollberger with the #gganimate package") +
    theme_minimal() +
    transition_states(state, transition_length = 1, state_length = 1)
```

