---
title: My First gganimate Plot
author: Derek Sollberger
date: '2019-01-08'
slug: my-first-gganimate-plot
categories:
  - stochastic processes
tags:
  - gganimate
header:
  caption: ''
  image: ''
---

Today I finally read a [tutorial](https://gganimate.com/articles/gganimate.html) on `gganimate`, and here I will build my first example.  My hope is to eventually simulate traversal on a directed graph.

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gganimate)
```

# Create Node Data

For this example, I will have 3 nodes (located at 3 vertices of a triangle), 100 dots distributed among the nodes, and a new placement over 21 states in time.

```{r}
N <- 100 # number of dots
T <- 20 # number of iterations

node_names <- c("Node 1", "Node 2", "Node 3")
iter <- rep(0, N)
node <- rep("Node 1", N)
x <- 10 + rnorm(N, 0, 2)
y <- 20 + rnorm(N, 0, 2)
states <- data.frame(iter, node, x, y)

for(t in 1:T){
  iter <- rep(t, N)
  node <- sample(node_names, N, replace = TRUE)
  x <- case_when(
    node == "Node 1" ~ 10 + rnorm(N, 0, 2),
    node == "Node 2" ~ 20 + rnorm(N, 0, 2),
    TRUE             ~ 30 + rnorm(N, 0, 2)
  )
  y <- case_when(
    node == "Node 1" ~ 20 + rnorm(N, 0, 2),
    node == "Node 2" ~ 30 + rnorm(N, 0, 2),
    TRUE             ~ 10 + rnorm(N, 0, 2)
  )
  this_state <- data.frame(iter, node, x, y)
  
  states <- rbind(states, this_state)
}
```

# Plot Network

Having made a new set of data at each state, the main tool is to use `transition_states` over the iterations.

```{r}
states %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point(aes(color = node), alpha = 0.33) +
  labs(title = "Draft of Network Animation",
       subtitle = "Iteration: {closest_state} of 20",
       caption = "Simulation of random placement") +
  theme_minimal() +
  transition_states(iter, transition_length = 1, state_length = 1)
```