---
title: Troubleshooting for R, RStudio, and Tidyverse installation
author: Derek Sollberger
date: '2020-01-29'
slug: r-troubleshooting-for-students
categories:
  - troubleshooting
tags:
  - r
  - rstudio
  - tidyverse
header:
  caption: ''
  image: ''
output:
  blogdown::html_page:
    toc: true
---

In this site, I hope to address most of the issues that appear when we try to install `R`, `RStudio`, and `tidyverse` in the classroom setting.

```{r, echo = FALSE}
# update to current version
R_version <- "3.6.2"
RStudio_version <- "1.2.5033"
```

# Why doesn't my R look like your R [software]?

There are two programs to install: `R`, then `RStudio`.  Please follow the installation instructions at https://dsollberger.netlify.com/post/r-installation-for-students/


# RStudio says that it cannot find the R binaries.

There are two programs to install: `R`, then `RStudio`.  Please follow the installation instructions at https://dsollberger.netlify.com/post/r-installation-for-students/


# We cannot install software (on iPad, Chromebook, etc.)

There is a fantastic alternative at [RStudio Cloud](https://rstudio.cloud/).  That is, one can do their R programming online.  There will be a few more steps involved compared to what we do in `RStudio` Desktop, but past students have gotten `RStudio Cloud` to work.


# tidyverse cannot be found

We need to install code packages before using code packages.  Run `install.packages("tidyverse")` in the console (lower-left area).  The installation will take up to 5 minutes.  This has to be done only once during the semester.


# [package] is not available for R version ...

To ensure compatibility between `R` and code packages, the R community almost always asks users to use the newest version of `R`.  Please install the newest versions of `R` and `RStudio`.  This semester we are using `R` version `r R_version` and `RStudio` version `r RStudio_version`.


# there is no package 'rlang'

Run this command in the console

* `install.packages("dplyr")`

If a choice appears that says something like `Do you want to install from sources the package which needs compilation?.`, select "No".

Alternatively: https://stackoverflow.com/questions/49721356/trouble-downloading-rlang-with-tidyverse


# there is no package 'broom'

The explanation would be a deep dive into the notion of *dependencies* (beyond the scope of the course).  For now, we dive right into this case's dependencies.  Run these commands in the console.

If a choice appears that says something like `Do you want to install from sources the package which needs compilation?.`, select "No".

* `install.packages("backports")`
* then `install.packages("zeallot")`
* then `install.packages("broom")`
* then `install.packages("tidyverse")`

Sources:

* https://community.rstudio.com/t/fail-to-install-package-broom/11307/3
* https://community.rstudio.com/t/error-lazy-loading-failed-for-package-broom/50777

# rlang and/or broom still do not work

Let's try this.  `tidyverse` is a *set* of packages that are supposed to work together well and emulate programming practices from other programming languages.  In our introductory forays into `R` programming, we do not need every package in the `tidyverse`.  Therefore, the simplest work-around is to load the needed packages individually.

```{r, message = FALSE, warning = FALSE}
#### instead of library("tidyverse") ####
library("dplyr")    #for the pipe function %>% and other SQL commands
library("ggplot2")  #modern data visualization
#library("readr")    #to load CSV data files
#library("readxl")   #to load Excel data files
#library("stringr")  #manipulate strings and regular expressions
#library("tidyr")    #to reshape data frames with functions like gather or spread
```


# How do we know which is the RMD file and which is the HTML file?

It will be useful if you can see file types in your computer.

* instructions for [Windows](http://kb.winzip.com/kb/entry/26/)
* instructions for [Mac OS X](https://www.idownloadblog.com/2014/10/29/how-to-show-or-hide-filename-extensions-in-os-x-yosemite/)

Also, the HTML file will probably have a similar icon as your web browser (Chrome, Firefox, etc.)

