---
title: Supervised Learning
author: Derek Sollberger
date: '2020-05-04'
slug: supervised-learning
categories:
  - machine learning
tags:
  - tidymodels
header:
  caption: ''
  image: ''
  preview: yes
---

# Supervised Learning

* **Supervised learning** has the goal of making predictions with a set of *known* labels for the response variable.
* In **unsupervised learning**, we try to find structure in the data of the response variable without predetermined labels.

Goal: predict the personality type of each character in Animal Crossing

# Data set: Animal Crossing

Source:  VillagerDB, MetaCritic, and TidyTuesday

* [Animal Crossing Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-05-05/readme.md)

```{r, message = FALSE, warning = FALSE}
library("caret")
library("randomForest")
library("tidymodels")
library("tidyverse")

# critic <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/critic.tsv')
# user_reviews <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/user_reviews.tsv')
# items <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/items.csv')
villagers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/villagers.csv')
```

## Exploratory Data Analysis

```{r}
dim(villagers)
```

```{r}
str(villagers)
#table(villagers$name)
#length(unique(villagers$name))
```

```{r}
table(villagers$personality)
```

## Personality across Gender

```{r}
villagers %>%
  ggplot(aes(x = personality, fill = gender)) +
  geom_bar(stat = "count", position = "dodge") +
  labs(title = "Getting to Know the Animal Crossing Villagers",
       subtitle = "and judging their personalities",
       caption = "Source: VillagerDB") +
  theme_minimal()
```

## Personality across Species

```{r}
villagers %>%
  ggplot(aes(x = personality, fill = species)) +
  geom_bar(stat = "count", position = "stack") +
  labs(title = "Getting to Know the Animal Crossing Villagers",
       subtitle = "and judging their personalities",
       caption = "Source: VillagerDB") +
  theme_minimal()
```


# Predictor Variables

```{r}
table(villagers$gender)
```
```{r}
table(villagers$species)
```

## Extracting Birth Month

```{r}
villagers <- villagers %>%
  separate(birthday, c("birth_month", "birth_day"), remove = FALSE)

villagers$birth_month_factor <- factor(villagers$birth_month, 
                                levels = 1:12)

table(villagers$birth_month_factor)
```

```{r}
villagers %>%
  ggplot(aes(x = personality, fill = birth_month_factor)) +
  geom_bar(color = "black", stat = "count", position = "stack") +
  labs(title = "Getting to Know the Animal Crossing Villagers",
       subtitle = "and judging their personalities",
       caption = "Source: VillagerDB") +
  theme_minimal()
```

```{r}
villagers %>%
  ggplot(aes(x = personality, fill = birth_month_factor)) +
  geom_bar(color = "black", stat = "count", position = "fill") +
  labs(title = "Getting to Know the Animal Crossing Villagers",
       subtitle = "and judging their personalities",
       caption = "Source: VillagerDB",
       y = "proportion") +
  scale_fill_manual(name = "Month of Birth",
                    labels = c("January", "February", "March", "April",
                                 "May", "June", "July", "August",
                                 "September", "October", "November", "December"),
                    values=sample(c("#4b48c9", "#5c90a8", "#86d5fe", "#c0d1ef",
                             "#edd1a2", "#948572", "#f8c79e", "#eec688",
                             "#642200", "#764936", "#fd974d", "#7b4a3c"))) +
  theme_minimal()
```

(color palette generated at [CSS Drive](http://www.cssdrive.com/imagepalette/index.php))

## model formula

* response variable: `personality`
* predictor variables: `gender`, `species`, `birth_month`
* model formula: `personality ~ gender + species + birth_month`

```{r}
predictor_variables <- c("gender", "species", "birth_month")
model_formula <- paste("personality~", paste(sprintf("`%s`", predictor_variables), collapse="+"))
```



# Data Split

```{r}
villagers_split <- initial_split(villagers)
villagers_train <- training(villagers_split)
villagers_test  <- testing(villagers_split)
```


# Random Forests

"Random forest models are ensembles of decision trees. A large number of decision tree models are created for the ensemble based on slightly different versions of the training set. When creating the individual decision trees, the fitting process encourages them to be as diverse as possible. The collection of trees are combined into the random forest model and, when a new sample is predicted, the votes from each tree are used to calculate the final predicted value for the new sample."
---[tidymodels.org](https://www.tidymodels.org/start/resampling/)

## Define the Forest

```{r}
random_forest_model <- 
  rand_forest(trees = 1000) %>% 
  set_engine("ranger") %>% 
  set_mode("classification")
```

## Fitting the Forest

```{r, eval = FALSE}
# CAUTION: at the time of filming, this code chunk did not work (model formula parser threw an error)
random_forest_fit <-
  random_forest_model %>%
  fit(personality ~ gender + species + birth_month, data = villagers_train)

random_forest_fit
```


## Visualizing the Forest

```{r, cache = TRUE}
# OLD-FASHIONED WAY with the caret package
model_rf <- caret::train(personality ~ gender + species + birth_month,
                         data = villagers_train, method = "rf")
model_rf

model_rpart <- caret::train(personality ~ gender + species + birth_month,
                         data = villagers_train, method = "rpart")
model_rpart
```

```{r, cache = TRUE, message = FALSE, warning = FALSE}
#source:  https://shiring.github.io/machine_learning/2017/03/16/rf_plot_ggraph

library("ggraph")
library("igraph")

tree_func <- function(final_model, 
                      tree_num) {
  
  # get tree by index
  tree <- randomForest::getTree(final_model, 
                                k = tree_num, 
                                labelVar = TRUE) %>%
    tibble::rownames_to_column() %>%
    # make leaf split points to NA, so the 0s won't get plotted
    mutate(`split point` = ifelse(is.na(prediction), `split point`, NA))
  
  # prepare data frame for graph
  graph_frame <- data.frame(from = rep(tree$rowname, 2),
                            to = c(tree$`left daughter`, tree$`right daughter`))
  
  # convert to graph and delete the last node that we don't want to plot
  graph <- graph_from_data_frame(graph_frame) %>%
    delete_vertices("0")
  
  # set node labels
  V(graph)$node_label <- gsub("_", " ", as.character(tree$`split var`))
  V(graph)$leaf_label <- as.character(tree$prediction)
  V(graph)$split <- as.character(round(tree$`split point`, digits = 2))
  
  # plot
  plot <- ggraph(graph, 'dendrogram') + 
    theme_bw() +
    geom_edge_link() +
    geom_node_point() +
    geom_node_text(aes(label = node_label), na.rm = TRUE, repel = TRUE) +
    geom_node_label(aes(label = split), vjust = 2.5, na.rm = TRUE, fill = "white") +
    geom_node_label(aes(label = leaf_label, fill = leaf_label), na.rm = TRUE, 
					repel = TRUE, colour = "white", fontface = "bold", show.legend = FALSE) +
    theme(panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          panel.background = element_blank(),
          plot.background = element_rect(fill = "white"),
          panel.border = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(size = 18))
  
  print(plot)
}

tree_num <- which(model_rf$finalModel$forest$ndbigtree == min(model_rf$finalModel$forest$ndbigtree))

tree_func(final_model = model_rf$finalModel, tree_num)
```


# Predictions

```{r}
predictions <- predict(model_rf, newdata = villagers_test)
confusionMatrix(predictions, factor(villagers_test$personality))
```

