---
title: 'TidyTuesday: Roman Emperors'
author: Derek Sollberger
date: '2020-06-19'
slug: tidytuesday-roman-emperors
categories:
  - tidyTuesday
tags:
  - data mining
  - ggplot
header:
  caption: ''
  image: ''
  preview: yes
---

# Introduction

Today, for practice with `ggplot2`, I wish to replicate @JoshuaFeldman's wonderful #TidyTuesday submission about the dataset of Roman emperors.

```{r, message = FALSE, warning = FALSE}
library("tidyverse")
```

* [TidyTuesday's Roman Emperor dataset](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-08-13) --- posted on August 13, 2019

```{r, message = FALSE, warning = FALSE}
# TidyTuesday's given line of code to load the data
emperors <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-13/emperors.csv")
```

# Exploring the Data

```{r}
dim(emperors)
colnames(emperors)
```

```{r}
emperors %>%
  filter(birth_prv != "Unknown") %>%
  ggplot(aes(x = birth_prv)) +
  geom_bar(aes(fill = birth_prv), stat = "count") +
  labs(title = "Birth Provinces of Roman Emperors",
       subtitle = "Years 62 BC to 347 AD",
       caption = "Data Source: TidyTuesday",
       x = "birth province",
       y = "number of emperors") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none")
```

```{r}
emperors %>%
  ggplot(aes(x = fct_rev(rise))) +
  geom_bar(aes(fill = rise), stat = "count") +
  coord_flip() +
  labs(title = "How Roman Emperors Rose to Power",
       subtitle = "Years 26 BC to 379 AD",
       caption = "Data Source: TidyTuesday",
       x = "",
       y = "number of emperors") +
  theme(legend.position = "none")
```

```{r}
emperors %>%
  filter(cause != "Unknown") %>%
  ggplot(aes(x = fct_rev(cause))) +
  geom_bar(aes(fill = cause), stat = "count") +
  coord_flip() +
  labs(title = "Causes of Death of Roman Emperors",
       subtitle = "Years 14 AD to 395 AD",
       caption = "Data Source: TidyTuesday",
       x = "",
       y = "number of emperors") +
  theme(legend.position = "none")
```

```{r}
emperors$dynasty_factor <- factor(emperors$dynasty,
                                  levels = unique(emperors$dynasty))

emperors %>%
  ggplot(aes(x = fct_rev(dynasty_factor))) +
  geom_bar(aes(fill = dynasty_factor), stat = "count") +
  coord_flip() +
  labs(title = "Roman Dynasties",
       subtitle = "Years 26 BC to 395 AD",
       caption = "Data Source: TidyTuesday",
       x = "",
       y = "number of emperors") +
  theme(legend.position = "none")
```

# Data Wrangling

```{r}
#extracting birth and death years
emperors <- emperors %>%
  separate(birth, 
           sep = "-", 
           into = c("birth_year", "birth_month", "birth_day"),
           remove = FALSE) %>%
  separate(death, 
           sep = "-", 
           into = c("death_year", "death_month", "death_day"),
           remove = FALSE)
```

```{r}
# remove leading zeros and convert from character to numeric
emperors$birth_year  <- as.numeric(emperors$birth_year)
emperors$birth_month <- as.numeric(emperors$birth_month)
emperors$birth_day   <- as.numeric(emperors$birth_day)
emperors$death_year  <- as.numeric(emperors$death_year)
emperors$death_month <- as.numeric(emperors$death_month)
emperors$death_day   <- as.numeric(emperors$death_day)
```

```{r}
# ensuring order of months for the graphs
emperors$birth_month <- factor(emperors$birth_month,
                               levels = 1:12)
emperors$death_month <- factor(emperors$death_month,
                               levels = 1:12)
```

```{r}
emperors %>%
  filter(!is.na(birth_month)) %>%
  ggplot(aes(x = birth_month)) +
  geom_bar(aes(fill = birth_month), stat = "count") +
  labs(title = "Birth Months of Roman Emperors",
       subtitle = "Years 26 BC to 347 AD",
       caption = "Data Source: TidyTuesday",
       x = "",
       y = "number of emperors") +
  theme(legend.position = "none")
```

```{r}
emperors %>%
  filter(!is.na(death_month)) %>%
  ggplot(aes(x = death_month)) +
  geom_bar(aes(fill = death_month), stat = "count") +
  labs(title = "Death Months of Roman Emperors",
       subtitle = "Years 14 AD to 395 AD",
       caption = "Data Source: TidyTuesday",
       x = "",
       y = "number of emperors") +
  theme(legend.position = "none")
```

```{r}
# only a few year data points were BC, so here is a quick fix for "negative years"
emperors$birth_year[1:3] <- -emperors$birth_year[1:3]
```

The `emperors` dataset had some missing values for `birth_year`, so here is what I am going to due to *impute* the missing values:

* calculate the lifespan of the other emperors
* calculate the average lifespan
* if `birth_year` is missing, subtract `death_year` by the average lifespan

```{r}
emperors$lifespan <- emperors$death_year - emperors$birth_year
average_emperor_lifespan <- round(mean(emperors$lifespan, na.rm = TRUE))

for(i in 1:nrow(emperors)){
  if(is.na(emperors$birth_year[i])){
    emperors$birth_year[i] <- emperors$death_year[i] - average_emperor_lifespan
  }
}
```

```{r}
#later, ggplot had a minor issue with 7 (i.e. greater than 6) causes of death
emperors$cause[emperors$cause == "Unknown"] <- "Natural Causes"
```


# Main Graph

As inspired by @JoshuaFeldman, the goals for today's main graph include

* birth years by province
* death years by cause of death
* reign as a line segment
* segment color by dynasty
* also delineate the two eras

```{r}
#extracting reign years
emperors <- emperors %>%
  separate(reign_start, 
           sep = "-", 
           into = c("reign_start_year", "reign_start_month", "reign_start_day"),
           remove = FALSE) %>%
  separate(reign_end, 
           sep = "-", 
           into = c("reign_end_year", "reign_end_month", "reign_end_day"),
           remove = FALSE)

emperors$reign_start_year  <- as.numeric(emperors$reign_start_year)
emperors$reign_start_month <- as.numeric(emperors$reign_start_month)
emperors$reign_start_day   <- as.numeric(emperors$reign_start_day)
emperors$reign_end_year    <- as.numeric(emperors$reign_end_year)
emperors$reign_end_month   <- as.numeric(emperors$reign_end_month)
emperors$reign_end_day     <- as.numeric(emperors$reign_end_day)

# only a few year data points were BC, so here is a quick fix for "negative years"
emperors$reign_start_year[1:2] <- -emperors$reign_start_year[1:2]
```

```{r}
main_plot <- emperors %>%
  ggplot() +
  geom_segment(aes(x = reign_start_year, y = reorder(name, -reign_start_year), 
                   xend = reign_end_year, yend = reorder(name, -reign_start_year),
                   color = dynasty_factor, group = dynasty_factor),
               data = emperors, size = 2) +
  geom_point(aes(x = birth_year, y = reorder(name, -reign_start_year),
                 group = birth_prv), 
             color = "black", size = 1) +
  geom_point(aes(x = death_year, y = reorder(name, -reign_start_year),
                 shape = cause, group = cause), 
             color = "black", size = 3) +
  #geom_label(aes(x = birth_year, y = reorder(name, -reign_start_year), label = name)) +
  scale_color_discrete(name = "Dynasty") +
  scale_shape_discrete(name = "Cause of Death") +
  labs(title = "Roman Emperors",
       subtitle = "Ordered by the start of their reign (small dots are birth years)",
       caption = "Data Source: TidyTuesday",
       x = "year",
       y = "") +
  theme(legend.position = "bottom")
```

```{r, fig.height = 8}
#plot
main_plot
```








