---
title: Duolingo Leagues
author: Derek Sollberger
date: '2020-01-13'
slug: duolingo-leagues
categories:
  - stochastic processes
tags:
  - math
header:
  caption: ''
  image: ''
---

`Duolingo`, the language learning app, places users in groups of 50 and assigns a `league` to each user to encourage competition.  The leagues are

* Bronze, Silver, Gold, Sapphire, Ruby, Emerald, Amethyst, Pearl, Obsidian, and Diamond (in that increasing order)

# What proportion of Duolingo users are in each league?

The rules are

* everyone starts in the Bronze League
* the top 15 percent of each group gets promoted to the next league up (measured weekly)
* the bottom 10 percent of each group is related downward

In this post, I will try out some stochastic processes calculations to answer that question.

```{r}
leagues <- c("Bronze", "Silver", "Gold", "Sapphire", "Ruby",
             "Emerald", "Amethyst", "Pearl", "Obsidian", "Diamond")

transition_matrix <- matrix( rep(0, 100), 10)

# trying shortcuts
diag(transition_matrix[-10,-1]) <- 15 #top 15 percent of each group gets promoted
diag(transition_matrix[-1,-10]) <- 10 #bottom 10 percent of each group is relegated
diag(transition_matrix)         <- 75 #the rest stay where they are

# fix the corners
transition_matrix[1,1]    <- 85
transition_matrix[10, 10] <- 85

# make row-stochastic (i.e. so each row adds up to one)
transition_matrix <- transition_matrix/100

# R allows user to label rows and columns
rownames(transition_matrix) <- leagues
colnames(transition_matrix) <- leagues

print(transition_matrix)
```

Knowing some about stochastic processes, we can either implement an initial distribution and employ matrix-vector multiplication over many iterations, or we can find an eigenvector.

```{r}
findStatDist <- function(P){
  # This function will compute the stationary distribution for a transition matrix
  # Input: row-stochastic matrix P
  # Output: row vector pi_vec
  
  # obtain left-eigenvector for lambda = 1
  x <- eigen(t(P))$vectors[,1] 
  
  # normalize the eigenvector in the one-norm
  pi_vec <- x / sum(x)
  pi_vec #return this vector
}

answer <- as.data.frame(round(100*findStatDist(transition_matrix)))

# R allows user to label rows and columns
rownames(answer) <- leagues
colnames(answer) <- "percentage"

print(answer)
```

