---
title: Analysis of NYT Sample of Covid-19 Obituaries
author: Derek Sollberger
date: '2020-05-24'
slug: analysis-of-nyt-sample-of-covid-19-obituaries
categories:
  - data mining
tags:
  - stringr
header:
  caption: ''
  image: ''
  preview: yes
---

# Obtaining the Data

On May 24, 2020, the *New York Times* published an article called "[An Incalcuable Loss](https://www.nytimes.com/interactive/2020/05/24/us/us-coronavirus-deaths-100000.html)"

```{}
"America is fast approaching a grim milestone in the coronavirus outbreak — each figure here represents one of the nearly 100,000 lives lost so far. But a count reveals only so much. Memories, gathered from obituaries across the country, help us to reckon with what was lost."
```

What I am trying to do today is summarize the 1001 obituaries.

```{r setup, include=FALSE}
library("ggridges")
library("tidyverse")
```

I started out by manually

* going to [https://www.nytimes.com/interactive/2020/05/24/us/us-coronavirus-deaths-100000.html](https://www.nytimes.com/interactive/2020/05/24/us/us-coronavirus-deaths-100000.html)
* right-clicking and selecting "View Source"
* copy-and-pasting the HTML code to a text document
* paring down the HTML code to focus on the obituaries.

```{}
<div id="g-obits">
	<div id="g-obit-frame">
		
			<div id="g-0" class="g-obit" data-date="February 6">
				<div class="g-obit-inner">
					<b>Auditor in Silicon Valley.</b>
					...
	...				
```

```{r}
texts <- readLines("obituaries.txt")
```

---

# Word Cloud

For one view, I wanted to make a word cloud to reveal the most commonly used words in the one-sentence obituaries.  I used the `stringr` package to extract the text in between the **bold** HTML tags (as well as removing the period).

```{r}
obits <- str_extract_all(texts, regex("(?<=<b>)(.*?)(?=.</b>)", multiline = TRUE))
```

There is `R` code and packages for making word clouds, but I find it best to use the site [wordclouds.com](https://www.wordclouds.com/) since that application has many neat features.  Thus, all I needed at this point is to save the results to upload to that site.

```{r}
write_lines(unlist(obits), "clean.txt")
```

---

# Density Plot of Ages

Q: Does Covid-19 only affect older people?

For this analysis, I seek the two- or three-digit numbers in between the *italics* HTML tags.

```{r}
ages <- as.integer( unlist( str_extract_all(texts, regex("(?<=,)(.[0-9]{2,3}?)(?=,)", multiline = TRUE) )))

summary(ages)
```

```{r}
data.frame(ages) %>%
  ggplot(aes(x = ages)) +
  geom_density(fill = "black") +
  labs(title = "Ages of Covid-19 Casualties",
       subtitle = "Sample of Obituaries from February to May 2020",
       caption = "Source: New York Times",
       x = "Age",
       y = "density") +
  theme(axis.text.y = element_blank())
```

---

# Spread by Week

Out of curiosity, here I hope to build a boxplot of the ages, but grouped by week of the year.

```{r}
dates <- unlist( str_extract_all(texts, regex("(?<=date=\")(.*?)(?=\")", multiline = TRUE) ))


# There was one missing age value in this data, so I am going to choose to delete the last date
# with hopes that the data still mostly lines up
dates <- dates[1:1000]

df <- data.frame(ages, dates)

# separate month and day
df <- df %>%
  separate(dates, into = c("month", "day"), remove = FALSE)

# Soon, I want to know the day of the year (i.e. how far into the year 2020), so here I make offsets,
df <- df %>%
  mutate(offset = case_when(month == "February" ~ 31,
                            month == "March" ~ 60,
                            month == "April" ~ 91,
                            TRUE ~ 121))

# Adding the offset and the obituary day to get the day of the year
df$day <- as.integer(df$day)
df <- df %>%
  mutate(day_of_year = offset + day)

# Finally, we divide by 7 to get the week of the year
df <- df %>%
  mutate(week_of_year = ceiling(day_of_year / 7))
```

```{r}
df %>%
  ggplot(aes(x = week_of_year, y = ages, fill = week_of_year, group = week_of_year)) +
  geom_boxplot() +
  labs(title = "US Covid-19 Deaths",
       subtitle = "Age Distributions by week, Sample of n = 1000 obituaries",
       caption = "Data Source: New York Times",
       x = "Week of Year (since start of 2020)",
       y = "Age") +
  theme(legend.position = "none")
```






