---
title: R Installation
author: Derek Sollberger
date: '2020-01-19'
slug: r-installation-for-students
categories:
  - teaching
tags:
  - workshop
header:
  caption: ''
  image: ''
---

```{r, echo = FALSE}
# update to current version
R_version <- "3.6.2"
RStudio_version <- "1.2.5033"
```

If you are planning to do the `R` assignments on your own computer (recommended), then here is a quick outline for obtaining the software.

* *There are two separate software programs*.  Most people find it easier to use `RStudio`. than just `R`, but you need to install `R` first before installing `RStudio` (analogously speaking:  you need an cell phone before you can use an cell phone case).
* *If you have R and RStudio from a previous course, you still need to update to the current versions!* (version `r R_version` of R and version `r RStudio_version` of RStudio)

# Downloading R

Go to [The Comprehensive R Archive Network](https://cran.r-project.org/) (CRAN)

Windows:

* click "Download R for Windows"
* click "install R for the first time"
* click "`r paste("Download R", R_version, "for Windows")`"
* click "Save File"

Mac: 

* click "Download R for (Mac) OS X"
* click "`r paste0("R-", R_version, ".pkg")`"
* click "Ok"

# Installing R

* Windows:  double-check icon for `r paste0("R-", R_version, "-win.exe")`
* Mac:  double-click icon for `r paste0("R-", R_version, ".pkg")`

For the rest of these steps, it is okay to accept the default selections (but I will outline them here for posterity):

* give computer permission to install
* select language
* agree to GNU General Public License
* select destination location
* select components
* Startup options: select "No (accept defaults)"
* select Start Menu Folder
* select Additional Tasks
* (wait a minute or two for installation to take place)
* "Click Finish to exit Setup"


# Downloading RStudio

Go to the [RStudio Desktop download page](https://rstudio.com/products/rstudio/download/#download) (i.e. we are using the *free*, desktop version of RStudio)

* Windows: click `r paste0("RStudio-", RStudio_version, ".exe")`, click "Save File"
* Mac: click `r paste0("RStudio-", RStudio_version, ".dmg")`, click "Ok"


# Installing RStudio

* Windows:  double-check icon for `r paste0("RStudio-", RStudio_version, ".exe")`
* Mac:  double-click icon for `r paste0("RStudio-", RStudio_version, ".dmg")`

For the rest of these steps, it is okay to accept the default selections (but I will outline them here for posterity):

* give computer permission to install
* "Click Next to continue"
* Choose Install Location
* Choose Start Menu Folder
* (wait a minute or two for installation to take place)
* "Click Finish to close Setup"


# Code Packages

For R coding, we will often need more code---called code packages. To ease our coding sessions, please run the following lines of code before class (in the lower-left `console` area of RStudio, just has to be done once this semester).

```{r, eval = FALSE}
package_list <- c("tidyverse", "readxl", "mosaic", "titanic")
install.packages(package_list)
```


# File Types

Also, while we are learning how I want homework assignments to be submitted, it will be useful if you can see file types in your computer.

* instructions for [Windows](http://kb.winzip.com/kb/entry/26/)
* instructions for [Mac OS X](https://www.idownloadblog.com/2014/10/29/how-to-show-or-hide-filename-extensions-in-os-x-yosemite/)


