---
title: Gender of Frasier Characters
author: Derek Sollberger
date: '2020-01-09'
slug: gender-of-frasier-characters
categories:
  - tidyverse
tags:
  - data mining
header:
  caption: ''
  image: ''
---

"I am not a man..."

For work, I need to take a list of names and try to infer the gender.  Here I try an R package on the character names in the TV show *Fraiser*.

# The gender package

```{r}
#install.packages("gender") #works fine

## user needs to download database too
#install_genderdata_package() #did not work ("error reading from connection")

## as suggested by the bug report at https://github.com/ropensci/drat/issues/6
#install.packages("devtools")
#library(devtools)
#devtools::install_github("ropensci/genderdata")
```

# Trial Run

```{r, message = FALSE, warning = FALSE}
library(gender)
library(ggpubr)
library(tidyverse)

gender("frasier", method = "ssa", years = c(1940, 1990))
```

# Cast of Characters

Now I will try to run the `gender` function over a list of names (criteria: characters that appeared in at least 6 episodes).

```{r}
#making a data frame of characters (i.e. the TV show characters)
name <- c("frasier", "daphne", "niles", "roz", "martin", 
          "eddie", "bulldog", "kenny", "gil", "noel", 
          "gertrude", "donny", "lilith", "bebe", "mel",
          "ronee", "alice", "julia", "frederick", "simon",
          "lana", "sherry", "kirby", "charlotte")
start_year <- 1940
end_year <- 1990
df <- data.frame(name, start_year, end_year, stringsAsFactors=FALSE)
main_cast <- df %>% slice(1:5)

gender_df(df, method = "ssa", name_col = "name", year_col = c("start_year", "end_year"))
```

# Sorting the Data

```{r}
results <- gender_df(df, method = "ssa", name_col = "name", year_col = c("start_year", "end_year"))

results_classified <- results %>%
  select(name, proportion_male, gender) %>%
  arrange(desc(proportion_male)) 

ggtexttable(results_classified, rows = NULL)
```

Note that "Mel" is a female character in the TV show.  The calculations are tougher when the input is a nickname.

```{r}
# just the main cast
main_cast_classified <- gender_df(main_cast, 
          method = "ssa", name_col = "name", 
          year_col = c("start_year", "end_year")) %>%
  select(name, proportion_male, gender) %>%
  arrange(desc(proportion_male))

ggtexttable(main_cast_classified, rows = NULL)
```